---
title: "SE Shuffle - Dataset 2g"
output: html_notebook
---

In this script, we will be testing how well SE does at recovering individual networks with LIONESS using Dataset 2g - 50 subjects, 100 taxa compared to those that have had their data shuffled (1 or 5 subjects, 1,3,or 5 edges shuffles)

# 1. Load Library
*** 
Packages required to run the script
```{r}
library(tidyverse)
library(SpiecEasi)
library(here)
library(lme4)
library(Matrix)
library(tscount)
library(patchwork)
library(MASS)
library(matrixcalc)
library(gridExtra)
library(devtools)
library(miaSim)
library(reshape2)
library(ggpubr)
library(broom)
library(ggnewscale)
library(coin)
```


# 2. Original Data and Edges
```{r}
Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Res_100sp.rds"))
```


# 3. LIONESS on Shuffle Data

## 3.1 - High to low
Strong edges, one species set to 0

### 3.1.1 - N1 I3
```{r}
Shuffle_N1_I3 <- read.csv(here("Data","GLV_SimData","Dataset_2g","Shuffle_N1_I3.csv"))
Shuffle_N1_I3 <- Shuffle_N1_I3 %>% column_to_rownames("X")
```

Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(Shuffle_N1_I3),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(net_data))
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Shuffle_N1_I3_Ind <- list()

# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(Shuffle_N1_I3)) {
      samplename <- rownames(Shuffle_N1_I3[j, ])
      nsamples <- nrow(Shuffle_N1_I3)
      net_single <- spiec.easi(data.matrix(Shuffle_N1_I3[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Shuffle_N1_I3_Ind[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Shuffle_N1_I3_Ind[[net_name]]) <- rownames(Shuffle_N1_I3_Ind[[net_name]]) <- colnames(Shuffle_N1_I3)
    }

# Store the results 
Dataset2g_Res$Shuffle_N1_I3_Ind <- Shuffle_N1_I3_Ind
```

```{r}
saveRDS(Dataset2g_Res,here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
```



### 3.1.2 - N5 I3
```{r}
Shuffle_N5_I3 <- read.csv(here("Data","GLV_SimData","Dataset_2g","Shuffle_N5_I3.csv"))
Shuffle_N5_I3 <- Shuffle_N5_I3 %>% column_to_rownames("X")
```


Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(Shuffle_N5_I3),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(net_data))
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Shuffle_N5_I3_Ind <- list()

# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(Shuffle_N5_I3)) {
      samplename <- rownames(Shuffle_N5_I3[j, ])
      nsamples <- nrow(Shuffle_N5_I3)
      net_single <- spiec.easi(data.matrix(Shuffle_N5_I3[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Shuffle_N5_I3_Ind[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Shuffle_N5_I3_Ind[[net_name]]) <- rownames(Shuffle_N5_I3_Ind[[net_name]]) <- colnames(Shuffle_N5_I3)
    }

# Store the results 
Dataset2g_Res$Shuffle_N5_I3_Ind <- Shuffle_N5_I3_Ind
```


```{r}
saveRDS(Dataset2g_Res,here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
```





### 3.1.3 - N1 I5
```{r}
Shuffle_N1_I5 <- read.csv(here("Data","GLV_SimData","Dataset_2g","Shuffle_N1_I5.csv"))
Shuffle_N1_I5 <- Shuffle_N1_I5 %>% column_to_rownames("X")
```

Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(Shuffle_N1_I5),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(net_data))
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Shuffle_N1_I5_Ind <- list()

# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(Shuffle_N1_I5)) {
      samplename <- rownames(Shuffle_N1_I5[j, ])
      nsamples <- nrow(Shuffle_N1_I5)
      net_single <- spiec.easi(data.matrix(Shuffle_N1_I5[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Shuffle_N1_I5_Ind[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Shuffle_N1_I5_Ind[[net_name]]) <- rownames(Shuffle_N1_I5_Ind[[net_name]]) <- colnames(Shuffle_N1_I5)
    }

# Store the results 
Dataset2g_Res$Shuffle_N1_I5_Ind <- Shuffle_N1_I5_Ind
```

```{r}
saveRDS(Dataset2g_Res,here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
```


### 3.1.4 - N5 I5
```{r}
Shuffle_N5_I5 <- read.csv(here("Data","GLV_SimData","Dataset_2g","Shuffle_N5_I5.csv"))
Shuffle_N5_I5 <- Shuffle_N5_I5 %>% column_to_rownames("X")
```


Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(Shuffle_N5_I5),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(net_data))
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Shuffle_N5_I5_Ind <- list()

# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(Shuffle_N5_I5)) {
      samplename <- rownames(Shuffle_N5_I5[j, ])
      nsamples <- nrow(Shuffle_N5_I5)
      net_single <- spiec.easi(data.matrix(Shuffle_N5_I5[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Shuffle_N5_I5_Ind[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Shuffle_N5_I5_Ind[[net_name]]) <- rownames(Shuffle_N5_I5_Ind[[net_name]]) <- colnames(Shuffle_N5_I5)
    }

# Store the results 
Dataset2g_Res$Shuffle_N5_I5_Ind <- Shuffle_N5_I5_Ind
```


```{r}
saveRDS(Dataset2g_Res,here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
```




## 3.2 - Low to High

### 3.2.1 - N1 I3

```{r}
Shuffle_N1_I3low <- read.csv(here("Data","GLV_SimData","Dataset_2g","Shuffle_N1_I3_low.csv"))
Shuffle_N1_I3low <- Shuffle_N1_I3low %>% column_to_rownames("X")
```

Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(Shuffle_N1_I3low),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(net_data))
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Shuffle_N1_I3low_Ind <- list()

# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(Shuffle_N1_I3low)) {
      samplename <- rownames(Shuffle_N1_I3low[j, ])
      nsamples <- nrow(Shuffle_N1_I3low)
      net_single <- spiec.easi(data.matrix(Shuffle_N1_I3low[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Shuffle_N1_I3low_Ind[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Shuffle_N1_I3low_Ind[[net_name]]) <- rownames(Shuffle_N1_I3low_Ind[[net_name]]) <- colnames(Shuffle_N1_I3low)
    }

# Store the results 
Dataset2g_Res$Shuffle_N1_I3low_Ind <- Shuffle_N1_I3low_Ind
```

```{r}
saveRDS(Dataset2g_Res,here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
```



### 3.2.2 - N5 I3
```{r}
Shuffle_N5_I3low <- read.csv(here("Data","GLV_SimData","Dataset_2g","Shuffle_N5_I3_low.csv"))
Shuffle_N5_I3low <- Shuffle_N5_I3low %>% column_to_rownames("X")
```


Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(Shuffle_N5_I3low),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(net_data))
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Shuffle_N5_I3low_Ind <- list()

# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(Shuffle_N5_I3low)) {
      samplename <- rownames(Shuffle_N5_I3low[j, ])
      nsamples <- nrow(Shuffle_N5_I3low)
      net_single <- spiec.easi(data.matrix(Shuffle_N5_I3low[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Shuffle_N5_I3low_Ind[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Shuffle_N5_I3low_Ind[[net_name]]) <- rownames(Shuffle_N5_I3low_Ind[[net_name]]) <- colnames(Shuffle_N5_I3low)
    }

# Store the results 
Dataset2g_Res$Shuffle_N5_I3low_Ind <- Shuffle_N5_I3low_Ind
```


```{r}
saveRDS(Dataset2g_Res,here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
```



##########################Current start


### 3.2.3 - N1 I5

```{r}
Shuffle_N1_I5low <- read.csv(here("Data","GLV_SimData","Dataset_2g","Shuffle_N1_I5_low.csv"))
Shuffle_N1_I5low <- Shuffle_N1_I5low %>% column_to_rownames("X")
```

Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(Shuffle_N1_I5low),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(net_data))
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Shuffle_N1_I5low_Ind <- list()

# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(Shuffle_N1_I5low)) {
      samplename <- rownames(Shuffle_N1_I5low[j, ])
      nsamples <- nrow(Shuffle_N1_I5low)
      net_single <- spiec.easi(data.matrix(Shuffle_N1_I5low[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Shuffle_N1_I5low_Ind[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Shuffle_N1_I5low_Ind[[net_name]]) <- rownames(Shuffle_N1_I5low_Ind[[net_name]]) <- colnames(Shuffle_N1_I5low)
    }

# Store the results 
Dataset2g_Res$Shuffle_N1_I5low_Ind <- Shuffle_N1_I5low_Ind
```

```{r}
saveRDS(Dataset2g_Res,here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
```



### 3.2.4 - N5 I5
```{r}
Shuffle_N5_I5low <- read.csv(here("Data","GLV_SimData","Dataset_2g","Shuffle_N5_I5_low.csv"))
Shuffle_N5_I5low <- Shuffle_N5_I5low %>% column_to_rownames("X")
```


Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(Shuffle_N5_I5low),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(net_data))
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Shuffle_N5_I5low_Ind <- list()

# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(Shuffle_N5_I5low)) {
      samplename <- rownames(Shuffle_N5_I5low[j, ])
      nsamples <- nrow(Shuffle_N5_I5low)
      net_single <- spiec.easi(data.matrix(Shuffle_N5_I5low[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Shuffle_N5_I5low_Ind[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Shuffle_N5_I5low_Ind[[net_name]]) <- rownames(Shuffle_N5_I5low_Ind[[net_name]]) <- colnames(Shuffle_N5_I5low)
    }

# Store the results 
Dataset2g_Res$Shuffle_N5_I5low_Ind <- Shuffle_N5_I5low_Ind
```


```{r}
saveRDS(Dataset2g_Res,here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Shuffle.rds"))
```





# 4. Comparison
## 4.1 Unaltered
```{r}
# Create an empty edge list
##########################################
Ind_Edge_Unalt_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Ind_Nets_unaltered"]])) {
    cov_matrix <- Dataset2g_Res[["Ind_Nets_unaltered"]][[i]]
    network_name <- names(Dataset2g_Res[["Ind_Nets_unaltered"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Ind_Edge_Unalt_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Ind_Edge_Unalt_Table <- Ind_Edge_Unalt_Table
Merged_Edge_Table <- do.call(rbind, Ind_Edge_Unalt_Table)

# merge with edges
##########################################
Unaltered_edges <- Merged_Edge_Table
```




## 4.2 N1 I3
```{r}
# Create an empty edge list
##########################################
Shuffle_N1_I3_Ind_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Shuffle_N1_I3_Ind"]])) {
    cov_matrix <- Dataset2g_Res[["Shuffle_N1_I3_Ind"]][[i]]
    network_name <- names(Dataset2g_Res[["Shuffle_N1_I3_Ind"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Shuffle_N1_I3_Ind_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Shuffle_N1_I3_Ind_Table <- Shuffle_N1_I3_Ind_Table

Shuffle_N1_I3_Ind <- do.call(rbind, Shuffle_N1_I3_Ind_Table)
Shuffle_N1_I3_Ind <- Shuffle_N1_I3_Ind %>% 
  dplyr::rename("Altered_Edge_weight" = "Edge_weight")
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Shuffle_N1_I3_Ind, 
                       by = c("SampleID", "Source", "Sink"))


# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_N1_I3 <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))
Combined_data_N1_I3$N <- 1
Combined_data_N1_I3$I <- 3

write.csv(Combined_data_N1_I3, here("Output", "SE_Lionness", "ShuffleEdges_N1I3.csv"))
```


## 4.3 N5 I3
```{r}
# Create an empty edge list
##########################################
Shuffle_N5_I3_Ind_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Shuffle_N5_I3_Ind"]])) {
    cov_matrix <- Dataset2g_Res[["Shuffle_N5_I3_Ind"]][[i]]
    network_name <- names(Dataset2g_Res[["Shuffle_N5_I3_Ind"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Shuffle_N5_I3_Ind_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Shuffle_N5_I3_Ind_Table <- Shuffle_N5_I3_Ind_Table

Shuffle_N5_I3_Ind <- do.call(rbind, Shuffle_N5_I3_Ind_Table)
Shuffle_N5_I3_Ind <- Shuffle_N5_I3_Ind %>% 
  dplyr::rename("Altered_Edge_weight" = "Edge_weight")
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Shuffle_N5_I3_Ind, 
                       by = c("SampleID", "Source", "Sink"))


# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_N5_I3 <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))
Combined_data_N5_I3$N <- 5
Combined_data_N5_I3$I <- 3

write.csv(Combined_data_N5_I3, here("Output", "SE_Lionness", "ShuffleEdges_N5I3.csv"))
```


## 4.4 N1 I5
```{r}
# Create an empty edge list
##########################################
Shuffle_N1_I5_Ind_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Shuffle_N1_I5_Ind"]])) {
    cov_matrix <- Dataset2g_Res[["Shuffle_N1_I5_Ind"]][[i]]
    network_name <- names(Dataset2g_Res[["Shuffle_N1_I5_Ind"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Shuffle_N1_I5_Ind_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Shuffle_N1_I5_Ind_Table <- Shuffle_N1_I5_Ind_Table

Shuffle_N1_I5_Ind <- do.call(rbind, Shuffle_N1_I5_Ind_Table)
Shuffle_N1_I5_Ind <- Shuffle_N1_I5_Ind %>% 
  dplyr::rename("Altered_Edge_weight" = "Edge_weight")
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Shuffle_N1_I5_Ind, 
                       by = c("SampleID", "Source", "Sink"))


# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_N1_I5 <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))
Combined_data_N1_I5$N <- 1
Combined_data_N1_I5$I <- 5

write.csv(Combined_data_N1_I5, here("Output", "SE_Lionness", "ShuffleEdges_N1I5.csv"))
```



## 4.5 N5 I5
```{r}
# Create an empty edge list
##########################################
Shuffle_N5_I5_Ind_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Shuffle_N5_I5_Ind"]])) {
    cov_matrix <- Dataset2g_Res[["Shuffle_N5_I5_Ind"]][[i]]
    network_name <- names(Dataset2g_Res[["Shuffle_N5_I5_Ind"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Shuffle_N5_I5_Ind_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Shuffle_N5_I5_Ind_Table <- Shuffle_N5_I5_Ind_Table

Shuffle_N5_I5_Ind <- do.call(rbind, Shuffle_N5_I5_Ind_Table)
Shuffle_N5_I5_Ind <- Shuffle_N5_I5_Ind %>% 
  dplyr::rename("Altered_Edge_weight" = "Edge_weight")
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Shuffle_N5_I5_Ind, 
                       by = c("SampleID", "Source", "Sink"))


# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_N5_I5 <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))
Combined_data_N5_I5$N <- 5
Combined_data_N5_I5$I <- 5

write.csv(Combined_data_N5_I5, here("Output", "SE_Lionness", "ShuffleEdges_N5I5.csv"))
```











## 4.6 N1 I3 Low
```{r}
# Create an empty edge list
##########################################
Shuffle_N1_I3low_Ind_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Shuffle_N1_I3low_Ind"]])) {
    cov_matrix <- Dataset2g_Res[["Shuffle_N1_I3low_Ind"]][[i]]
    network_name <- names(Dataset2g_Res[["Shuffle_N1_I3low_Ind"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Shuffle_N1_I3low_Ind_Table[[edge_table_name]] <- adj
}

Dataset2g_Res$Shuffle_N1_I3low_Ind_Table <- Shuffle_N1_I3low_Ind_Table

Shuffle_N1_I3_Ind_low <- do.call(rbind, Shuffle_N1_I3low_Ind_Table)
Shuffle_N1_I3_Ind_low <- Shuffle_N1_I3_Ind_low %>% 
  dplyr::rename("Altered_Edge_weight" = "Edge_weight")
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Shuffle_N1_I3_Ind_low, 
                       by = c("SampleID", "Source", "Sink"))


# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_N1_I3_low <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))
Combined_data_N1_I3_low$N <- 1
Combined_data_N1_I3_low$I <- 3

write.csv(Combined_data_N1_I3_low, here("Output", "SE_Lionness", "ShuffleEdges_N1I3_low.csv"))
```


## 4.7 N5 I3 Low
```{r}
# Create an empty edge list
##########################################
Shuffle_N5_I3low_Ind_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Shuffle_N5_I3low_Ind"]])) {
    cov_matrix <- Dataset2g_Res[["Shuffle_N5_I3low_Ind"]][[i]]
    network_name <- names(Dataset2g_Res[["Shuffle_N5_I3low_Ind"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Shuffle_N5_I3low_Ind_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Shuffle_N5_I3low_Ind_Table <- Shuffle_N5_I3low_Ind_Table

Shuffle_N5_I3low_Ind <- do.call(rbind, Shuffle_N5_I3low_Ind_Table)
Shuffle_N5_I3low_Ind <- Shuffle_N5_I3low_Ind %>% 
  dplyr::rename("Altered_Edge_weight" = "Edge_weight")
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Shuffle_N5_I3low_Ind, 
                       by = c("SampleID", "Source", "Sink"))


# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_N5_I3low <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))
Combined_data_N5_I3low$N <- 5
Combined_data_N5_I3low$I <- 3

write.csv(Combined_data_N5_I3low, here("Output", "SE_Lionness", "ShuffleEdges_N5I3_low.csv"))
```


## 4.8 N1 I5 Low
```{r}
# Create an empty edge list
##########################################
Shuffle_N1_I5low_Ind_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Shuffle_N1_I5low_Ind"]])) {
    cov_matrix <- Dataset2g_Res[["Shuffle_N1_I5low_Ind"]][[i]]
    network_name <- names(Dataset2g_Res[["Shuffle_N1_I5low_Ind"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Shuffle_N1_I5low_Ind_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Shuffle_N1_I5low_Ind_Table <- Shuffle_N1_I5low_Ind_Table

Shuffle_N1_I5low_Ind <- do.call(rbind, Shuffle_N1_I5low_Ind_Table)
Shuffle_N1_I5low_Ind <- Shuffle_N1_I5low_Ind %>% 
  dplyr::rename("Altered_Edge_weight" = "Edge_weight")
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Shuffle_N1_I5low_Ind, 
                       by = c("SampleID", "Source", "Sink"))


# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_N1_I5low <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))
Combined_data_N1_I5low$N <- 1
Combined_data_N1_I5low$I <- 5

write.csv(Combined_data_N1_I5low, here("Output", "SE_Lionness", "ShuffleEdges_N1I5_low.csv"))
```



## 4.9 N5 I5 Low
```{r}
# Create an empty edge list
##########################################
Shuffle_N5_I5low_Ind_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Shuffle_N5_I5low_Ind"]])) {
    cov_matrix <- Dataset2g_Res[["Shuffle_N5_I5low_Ind"]][[i]]
    network_name <- names(Dataset2g_Res[["Shuffle_N5_I5low_Ind"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Shuffle_N5_I5low_Ind_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Shuffle_N5_I5low_Ind_Table <- Shuffle_N5_I5low_Ind_Table

Shuffle_N5_I5low_Ind <- do.call(rbind, Shuffle_N5_I5low_Ind_Table)
Shuffle_N5_I5low_Ind <- Shuffle_N5_I5low_Ind %>% 
  dplyr::rename("Altered_Edge_weight" = "Edge_weight")
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Shuffle_N5_I5low_Ind, 
                       by = c("SampleID", "Source", "Sink"))


# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_N5_I5low <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))
Combined_data_N5_I5low$N <- 5
Combined_data_N5_I5low$I <- 5

write.csv(Combined_data_N5_I5low, here("Output", "SE_Lionness", "ShuffleEdges_N5I5_low.csv"))
```









