---
title: "SE LIONESS Summary"
output: html_notebook
---

This script will summarize the the analysis of LIONESS and SE using the SE simulated data, Dataset 2g


# 1. Load Library
*** 
Packages required to run the script
```{r}
library(tidyverse)
library(SpiecEasi)
library(here)
library(lme4)
library(Matrix)
library(tscount)
library(patchwork)
library(MASS)
library(matrixcalc)
library(gridExtra)
library(devtools)
library(miaSim)
library(reshape2)
library(ggpubr)
library(broom)
library(ggnewscale)
library(coin)
```


# 2. Figure #1A

## 2.1 Data Prep
```{r}
# 50 Subjects 50 Taxa
######################################
Edges_50sbj50taxa <- read.csv(here("Output", "SE_Lionness", "Edges_50sbj50taxa.csv"))
Edges_50sbj50taxa$N_Subjects <- 50
Edges_50sbj50taxa$N_Species <- 50
Edges_50sbj50taxa$Subject_Alt <- "Sbj1"
Edges_50sbj50taxa$Species_Alt <- "sp28"

# Sort to top 10
Edges_50sbj50taxa <- Edges_50sbj50taxa %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 10)

# 100 Subjects 50 Taxa
######################################
Edges_100sbj <- read.csv(here("Output", "SE_Lionness", "Edges_100sbj.csv"))
Edges_100sbj$N_Subjects <- 100
Edges_100sbj$N_Species <- 50
Edges_100sbj$Subject_Alt <- "Sbj1"
Edges_100sbj$Species_Alt <- "sp28"

# Sort to top 10
Edges_100sbj <- Edges_100sbj %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 10)

# 50 Subjects 100 Taxa
######################################
Edges_100sp <- read.csv(here("Output", "SE_Lionness", "Edges_100sp.csv"))
Edges_100sp$N_Subjects <- 50
Edges_100sp$N_Species <- 100
Edges_100sp$Subject_Alt <- "Sbj1"
Edges_100sp$Species_Alt <- "sp65"

# Sort to top 10
Edges_100sp <- Edges_100sp %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 10)

# 100 Subjects 100 Taxa
######################################
Edges_100sp_100sbj <- read.csv(here("Output", "SE_Lionness", "Edges_100sbj100sp.csv"))
Edges_100sp_100sbj$N_Subjects <- 100
Edges_100sp_100sbj$N_Species <- 100
Edges_100sp_100sbj$Subject_Alt <- "Sbj1"
Edges_100sp_100sbj$Species_Alt <- "sp98"

# Sort to top 10
Edges_100sp_100sbj <- Edges_100sp_100sbj %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 10)
```

## 2.2 Plot the data
```{r}
# Merge Data
All_data <- rbind(Edges_50sbj50taxa, Edges_100sbj, Edges_100sp,Edges_100sp_100sbj)

# Add Plot Colors
All_data <- All_data %>% 
  mutate(Plot_Colors = case_when(
    SampleID == Subject_Alt & Source == Species_Alt ~ "Altered Subject & Species",
    SampleID == Subject_Alt & Sink == Species_Alt ~ "Altered Subject & Species",
    SampleID != Subject_Alt & Source == Species_Alt ~ "Altered Species",
    SampleID != Subject_Alt & Sink == Species_Alt ~ "Altered Species",
    SampleID == Subject_Alt & Source != Species_Alt ~ "Altered Subject",
    SampleID == Subject_Alt & Sink != Species_Alt ~ "Altered Subject",
    TRUE ~ "Neither Altered"))

# Make the Facet Plot

ggplot(All_data, aes(x = reorder(paste(Source, Sink, sep="-"), abs(Edge_diff)), 
                      y = abs(Edge_diff), fill = Plot_Colors)) +
  geom_col(color = "black") +
  facet_grid(N_Subjects ~ N_Species, scales = "free", space = "free_x", 
             labeller = labeller(N_Subjects = label_value, N_Species = label_value)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Edge", y = "Absolute Edge Difference") +
  scale_fill_manual(name = "Edge Type", 
                     values = c("Altered Subject & Species" = "darkgreen", 
                                "Altered Species" = "purple",
                                "Altered Subject" = "orange",
                                "Neither Altered" = "grey"))  +
  theme(axis.text.x = element_text(family = "sans", color = "black", size = 10),
        axis.text.y = element_text(family = "sans", color = "black", size = 14),
        axis.title.x = element_text(family = "sans", color = "black", size = 14),
        axis.title.y = element_text(family = "sans", color = "black", size = 14),
        strip.text.x = element_text(family = "sans", color = "black", size = 14),
        strip.text.y = element_text(family = "sans", color = "black", size = 14),
        legend.position = "none")
```


# 3. Figure #1B

## 3.1 Prep the data
Get the index list from the simulation code
```{r}
N1 <- "Sbj21_Time2"
N5 <- c("Sbj21_Time2","Sbj23_Time2","Sbj32_Time2",
        "Sbj34_Time2","Sbj30_Time2")

I3 <- data.frame(Index_Source = c("sp11", "sp30", "sp7"))
#I3 <- data.frame(Index_Source = c("sp11", "sp30", "sp7"), 
#                 Index_Sink = c("sp98", "sp65", "sp33"))

I5 <- data.frame(Index_Source = c("sp11", "sp30", "sp7", "sp54", "sp67"))

#I5 <- data.frame(Index_Source = c("sp11", "sp30", "sp7", "sp54", "sp67"), 
#                 Index_Sink = c("sp98", "sp65", "sp33", "sp97", "sp88"))
```




1 or 5 Subjects, 3 Indices changed
```{r}
# N1 I3
######################################################
ShuffleEdges_N1_I3 <- read.csv(here("Output", "SE_Lionness", "ShuffleEdges_N1I3.csv"))

ShuffleEdges_N1_I3 <- ShuffleEdges_N1_I3 %>%
  rowwise() %>% 
  mutate(Altered_Subject = if_else(SampleID %in% N1, "Altered", "Not-Altered"),
        Species_Altered = if_else(
          any(Source %in% I3$Index_Source | Sink %in% I3$Index_Source),
          "Altered", "Not-Altered")) %>%
  ungroup()

# Filter to Edges Changed
ShuffleEdges_N1_I3_filt <- ShuffleEdges_N1_I3 %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 25)

# N5 I3
######################################################

ShuffleEdges_N5_I3 <- read.csv(here("Output", "SE_Lionness", "ShuffleEdges_N5I3.csv"))

ShuffleEdges_N5_I3 <- ShuffleEdges_N5_I3 %>%
  rowwise() %>% 
  mutate(Altered_Subject = if_else(SampleID %in% N5, "Altered", "Not-Altered"),
        Species_Altered = if_else(
          any(Source %in% I3$Index_Source | Sink %in% I3$Index_Source),
          "Altered", "Not-Altered")) %>%
  ungroup()

# Filter to Edges Changed
ShuffleEdges_N5_I3_filt <- ShuffleEdges_N5_I3 %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 25)

```



1 or 5 Subjects, 5 Indices changed
```{r}
# N1 I5
######################################################
ShuffleEdges_N1_I5 <- read.csv(here("Output", "SE_Lionness", "ShuffleEdges_N1I5.csv"))


ShuffleEdges_N1_I5 <- ShuffleEdges_N1_I5 %>%
  rowwise() %>% 
  mutate(Altered_Subject = if_else(SampleID %in% N1, "Altered", "Not-Altered"),
        Species_Altered = if_else(
          any(Source %in% I5$Index_Source | Sink %in% I5$Index_Source),
          "Altered", "Not-Altered")) %>%
  ungroup()

# Filter to Edges Changed
ShuffleEdges_N1_I5_filt <- ShuffleEdges_N1_I5 %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 25)

# N5 I5
######################################################
ShuffleEdges_N5_I5 <- read.csv(here("Output", "SE_Lionness", "ShuffleEdges_N5I5.csv"))



ShuffleEdges_N5_I5 <- ShuffleEdges_N5_I5 %>%
  rowwise() %>% 
  mutate(Altered_Subject = if_else(SampleID %in% N5, "Altered", "Not-Altered"),
        Species_Altered = if_else(
          any(Source %in% I5$Index_Source | Sink %in% I5$Index_Source),
          "Altered", "Not-Altered")) %>%
  ungroup()

# Filter to Edges Changed
ShuffleEdges_N5_I5_filt <- ShuffleEdges_N5_I5 %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 25)
```


## 3.2 Plot the data
```{r}
# Merge Data
All_data <- rbind(ShuffleEdges_N1_I3_filt,ShuffleEdges_N5_I3_filt,
                  ShuffleEdges_N1_I5_filt,ShuffleEdges_N5_I5_filt)

# Add Plot Colors
All_data <- All_data %>% 
  mutate(Plot_Colors = case_when(
    Altered_Subject == "Altered" & 
    Species_Altered == "Altered" ~ "Altered Subject & Species",
        Altered_Subject != "Altered" & 
        Species_Altered == "Altered" ~ "Altered Species",
      Altered_Subject == "Altered" & 
      Species_Altered != "Altered" ~ "Altered Subject",
    TRUE ~ "Neither Altered")) %>% 
  rename("N" ="N_Subjects") %>% 
  rename("I" ="N_Species")

# Make the Facet Plot

ggplot(All_data, aes(x = reorder(paste(Source, Sink, sep="-"), abs(Edge_diff)), 
                      y = abs(Edge_diff), fill = Plot_Colors)) +
  geom_col(color = "black") +
  facet_grid(N_Subjects ~ N_Species, scales = "free", space = "free_x", 
             labeller = labeller(N_Subjects = label_value, N_Species = label_value)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Edge", y = "Absolute Edge Difference") +
  scale_fill_manual(name = "Edge Type", 
                     values = c("Altered Subject & Species" = "darkgreen", 
                                "Altered Species" = "purple",
                                "Altered Subject" = "orange",
                                "Neither Altered" = "grey"))  +
  theme(axis.text.x = element_text(family = "sans", color = "black", size = 6),
        axis.text.y = element_text(family = "sans", color = "black", size = 14),
        axis.title.x = element_text(family = "sans", color = "black", size = 14),
        axis.title.y = element_text(family = "sans", color = "black", size = 14),
        strip.text.x = element_text(family = "sans", color = "black", size = 14),
        strip.text.y = element_text(family = "sans", color = "black", size = 14),
        legend.position = "none")
```





# 4. Figure #1C
## 4.1 Prep the data
Get the index list from the simulation code
```{r}
N1 <- "Sbj21_Time2"
N5 <- c("Sbj21_Time2","Sbj23_Time2","Sbj32_Time2",
        "Sbj34_Time2","Sbj30_Time2")


I3_low <- data.frame(Index_Source = c("sp1", "sp3", "sp4"), 
                 Index_Sink = c("sp2", "sp5", "sp6"))

I5_low <- data.frame(Index_Source = c("sp1", "sp3", "sp4", "sp7", "sp9"), 
                 Index_Sink = c("sp2", "sp5", "sp6", "sp8", "sp10"))
```



1 or 5 Subjects, 3 Indices changed
```{r}
# N1 I3_low
######################################################
ShuffleEdges_N1_I3 <- read.csv(here("Output", "SE_Lionness", "ShuffleEdges_N1I3_low.csv"))

ShuffleEdges_N1_I3 <- ShuffleEdges_N1_I3 %>%
  rowwise() %>% 
  mutate(Altered_Subject = if_else(SampleID %in% N1, "Altered", "Not-Altered"),
        Both_Altered_Taxa = if_else(
          any(Source == I3_low$Index_Source & Sink == I3_low$Index_Sink), 
          "Altered", "Not-Altered"),
        One_Altered_Taxa = if_else(
          any(Source %in% I3_low$Index_Source | Sink %in% I3_low$Index_Sink | 
          Sink %in% I3_low$Index_Source | Source %in% I3_low$Index_Sink), 
          "Altered", "Not-Altered")) %>%
  ungroup()

# Filter to Edges Changed
ShuffleEdges_N1_I3_filt <- ShuffleEdges_N1_I3 %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 25)

# N5 I3_low
######################################################

ShuffleEdges_N5_I3 <- read.csv(here("Output", "SE_Lionness", "ShuffleEdges_N5I3_low.csv"))

ShuffleEdges_N5_I3 <- ShuffleEdges_N5_I3 %>%
  rowwise() %>% 
  mutate(Altered_Subject = if_else(SampleID %in% N5, "Altered", "Not-Altered"),
        Both_Altered_Taxa = if_else(
          any(Source == I3_low$Index_Source & Sink == I3_low$Index_Sink), 
          "Altered", "Not-Altered"),
        One_Altered_Taxa = if_else(
          any(Source %in% I3_low$Index_Source | Sink %in% I3_low$Index_Sink | 
          Sink %in% I3_low$Index_Source | Source %in% I3_low$Index_Sink), 
          "Altered", "Not-Altered")) %>%
  ungroup()

# Sort to top 15
ShuffleEdges_N5_I3_filt <- ShuffleEdges_N5_I3 %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 25)

```


1 or 5 Subjects, 5 Indices changed
```{r}
# N1 I3_low
######################################################
ShuffleEdges_N1_I5 <- read.csv(here("Output", "SE_Lionness", "ShuffleEdges_N1I5_low.csv"))

ShuffleEdges_N1_I5 <- ShuffleEdges_N1_I5 %>%
  rowwise() %>% 
  mutate(Altered_Subject = if_else(SampleID %in% N1, "Altered", "Not-Altered"),
        Both_Altered_Taxa = if_else(
          any(Source == I5_low$Index_Source & Sink == I5_low$Index_Sink), 
          "Altered", "Not-Altered"),
        One_Altered_Taxa = if_else(
          any(Source %in% I5_low$Index_Source | Sink %in% I5_low$Index_Sink | 
          Sink %in% I5_low$Index_Source | Source %in% I5_low$Index_Sink), 
          "Altered", "Not-Altered")) %>%
  ungroup()

# Filter to Edges Changed
ShuffleEdges_N1_I5_filt <- ShuffleEdges_N1_I5 %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 25)

# N5 I3_low
######################################################

ShuffleEdges_N5_I5 <- read.csv(here("Output", "SE_Lionness", "ShuffleEdges_N5I5_low.csv"))

ShuffleEdges_N5_I5 <- ShuffleEdges_N5_I5 %>%
  rowwise() %>% 
  mutate(Altered_Subject = if_else(SampleID %in% N5, "Altered", "Not-Altered"),
        Both_Altered_Taxa = if_else(
          any(Source == I5_low$Index_Source & Sink == I5_low$Index_Sink), 
          "Altered", "Not-Altered"),
        One_Altered_Taxa = if_else(
          any(Source %in% I5_low$Index_Source | Sink %in% I5_low$Index_Sink | 
          Sink %in% I5_low$Index_Source | Source %in% I5_low$Index_Sink), 
          "Altered", "Not-Altered")) %>%
  ungroup()

# Sort to top 25
ShuffleEdges_N5_I5_filt <- ShuffleEdges_N5_I5 %>%
  arrange(desc(abs(Edge_diff))) %>% slice_head(n = 25)

```


```{r}
# Merge the Data
Shuffle_merged <- rbind(ShuffleEdges_N1_I3_filt, ShuffleEdges_N5_I3_filt,
                        ShuffleEdges_N1_I5_filt, ShuffleEdges_N5_I5_filt)

Shuffle_merged <- Shuffle_merged %>% 
  mutate(Plot_Colors = case_when(
    Altered_Subject == "Altered" & Both_Altered_Taxa == "Altered" 
                                    ~ "Altered Subject & Species",
    Altered_Subject == "Altered" & One_Altered_Taxa == "Not-Altered" 
                                    ~ "Altered Subject",
    Altered_Subject == "Not-Altered" & One_Altered_Taxa == "Altered" 
                                    ~ "Altered Species",
    TRUE ~ "Neither Altered")) %>%
  mutate(I = case_when(I == "3" ~ 6,
                       I == "5" ~ 10)) %>%
  rename("N"="N_Subjects") %>%
  rename("I"="N_Species")

# Make the Facet Plot

ggplot(Shuffle_merged, 
       aes(x = reorder(paste(Source, Sink, sep="-"), abs(Edge_diff)), 
                      y = abs(Edge_diff), fill = Plot_Colors)) +
  geom_col(color = "black") +
  facet_grid(N_Subjects ~ N_Species, scales = "free", space = "free_x", 
             labeller = labeller(N_Subjects = label_value, N_Species = label_value)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Edge", y = "Absolute Edge Difference") +
  scale_fill_manual(name = "Edge Type", 
                     values = c("Altered Subject & Species" = "darkgreen", 
                                "Altered Species" = "purple",
                                "Altered Subject" = "orange",
                                "Neither Altered" = "grey"))  +
  theme(axis.text.x = element_text(family = "sans", color = "black", size = 8),
        axis.text.y = element_text(family = "sans", color = "black", size = 14),
        axis.title.x = element_text(family = "sans", color = "black", size = 14),
        axis.title.y = element_text(family = "sans", color = "black", size = 14),
        strip.text.x = element_text(family = "sans", color = "black", size = 14),
        strip.text.y = element_text(family = "sans", color = "black", size = 14),
        legend.position = "none")

```




