---
title: "SE & LIONNESS - Dataset 2g 50/100"
output: html_notebook
---


In this script, we will be testing how well SE does at recovering individual networks with LIONESS using Dataset 2g - 50 subjects, 100 taxa. 

# 1. Load Library
*** 
Packages required to run the script
```{r}
library(tidyverse)
library(SpiecEasi)
library(here)
library(lme4)
library(Matrix)
library(tscount)
library(patchwork)
library(MASS)
library(matrixcalc)
library(gridExtra)
library(devtools)
library(miaSim)
library(reshape2)
library(ggpubr)
library(broom)
library(ggnewscale)
library(coin)
```



# 2. Data Prep

Read in the data
```{r}
# Read in the raw data
original_data <- read.csv(here("Data", "GLV_SimData", "Dataset_2g",
                               "SE_50sbj100taxa.csv"))
original_data <- original_data %>% column_to_rownames("X") 
counts_data <- original_data

# Cor
original_cor <- read.csv(here("Data", "GLV_SimData", "Dataset_2g",
                               "Cor_50sbj100taxa.csv"))
```



Store the overall count table in a list that will be used downstream
```{r}
Dataset2g_Res <- list()
Dataset2g_Res$Counts <- counts_data
Dataset2g_Res$Cor <- original_cor
```


# 3. Individual changes

Here we will run SE and Lionness for all the data. Then ID which edge is the strongest in the overall SE network, and change those taxa in one individuals counts. We will then go back and run their samples through lionness and see if it can detect the change

## 3.1 - Group Network

Run SE to get a network for all the data

```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(counts_data),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )
  

# Store the results in Dataset1_Res
Dataset2g_Res$SpeicEasi_Time <- net_data
```

## 3.2 - Unaltered Ind Networks

Run LIONESS to get the individual networks of the unaltered  data
```{r}
# Prep the data
################################################
# List to store individual networks
Ind_Nets_unaltered <- list()

# Overall network
net_all <- as.matrix(SpiecEasi::getOptCov(Dataset2g_Res[["SpeicEasi_Time"]]))


# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(counts_data)) {
      samplename <- rownames(counts_data[j, ])
      nsamples <- nrow(counts_data)
      net_single <- spiec.easi(data.matrix(counts_data[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Ind_Nets_unaltered[[net_name]] <- 
        nsamples * (net_all - net_single_cov) + net_single_cov
      colnames(Ind_Nets_unaltered[[net_name]]) <- rownames(Ind_Nets_unaltered[[net_name]]) <- colnames(counts_data)
    }

# Store the results 
Dataset2g_Res$Ind_Nets_unaltered <- Ind_Nets_unaltered
```


## 3.3 - Find Edge
Pick up a strong edge
```{r}
# Extract the Covariance Matrix
net_single_cov <- as.matrix(getOptCov(Dataset2g_Res$SpeicEasi_Time))

# Set Diagnols to 0
diagonal_indices <- 1:nrow(net_single_cov)
net_single_cov[cbind(diagonal_indices, diagonal_indices)] <- 0

# Add species name
colnames(net_single_cov) <- rownames(net_single_cov) <- colnames(counts_data)

# Find the max value
max(net_single_cov, na.rm = TRUE)
which(net_single_cov == max(net_single_cov, na.rm = TRUE), arr.ind = TRUE)
```


## 3.4 - Change Values
Based on these, set species 65 to 0 in subject one
```{r}
counts_altered <- counts_data
counts_altered["Sbj1_Time2", 65] <- 0

```


## 3.5 - Individual Networks

Get the overall network first
```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(counts_altered),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )
  

# Store the results in Dataset1_Res
Dataset2g_Res$SpeicEasi_Time_alt <- net_data
```


Run LIONESS to get the individual networks
```{r}
# Prep the data
################################################
# List to store individual networks
Ind_Nets_alt <- list()

# Combined Network
net_all <- as.matrix(SpiecEasi::getOptCov(Dataset2g_Res[["SpeicEasi_Time_alt"]]))


# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(counts_altered)) {
      samplename <- rownames(counts_altered[j, ])
      nsamples <- nrow(counts_altered)
      net_single <- spiec.easi(data.matrix(counts_altered[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Ind_Nets_alt[[net_name]] <- nsamples * (net_all -
                                                       net_single_cov) + net_single_cov
      colnames(Ind_Nets_alt[[net_name]]) <- rownames(Ind_Nets_alt[[net_name]]) <- colnames(counts_altered)
    }

# Store the results 
Dataset2g_Res$Ind_Nets_alt <- Ind_Nets_alt
```



## 3.6 - Check Recovery

## 3.6.1 Unaltered Edges
```{r}
#Dataset2g_Res <- readRDS(here("Output", "SE_Lionness", "Dataset2g_Res_100sp.rds"))

# Create an empty edge list
##########################################
Ind_Edge_Unalt_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Ind_Nets_unaltered"]])) {
    cov_matrix <- Dataset2g_Res[["Ind_Nets_unaltered"]][[i]]
    network_name <- names(Dataset2g_Res[["Ind_Nets_unaltered"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Ind_Edge_Unalt_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Ind_Edge_Unalt_Table <- Ind_Edge_Unalt_Table
Merged_Edge_Table <- do.call(rbind, Ind_Edge_Unalt_Table)

# merge with edges
##########################################
Unaltered_edges <- Merged_Edge_Table
```


## 3.6.2 Altered Edges
Convert altered networks to an edge list
```{r}
# Create an empty edge list
##########################################
Ind_Edge_altered_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_Res[["Ind_Nets_alt"]])) {
    cov_matrix <- Dataset2g_Res[["Ind_Nets_alt"]][[i]]
    network_name <- names(Dataset2g_Res[["Ind_Nets_alt"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Ind_Edge_altered_Table[[edge_table_name]] <- adj
  }
Dataset2g_Res$Ind_Edge_altered_Table <- Ind_Edge_altered_Table
Merged_Edge_Table <- do.call(rbind, Ind_Edge_altered_Table)


# merge with sample data
##########################################
Altered_data <- Merged_Edge_Table
# rename Edge_weight to Altered_Edge_weight
Altered_data <- Altered_data %>% rename("Edge_weight"="Altered_Edge_weight")
```


Combine the data
```{r}
# Combine the data
#####################################
Combined_data <- merge(Unaltered_edges, Altered_data, 
                       by = c("SampleID", "Source", "Sink"))

Combined_data$SampleID <- gsub("_Time2", "", Combined_data$SampleID)

# create a difference column
Combined_data$Edge_diff <- Combined_data$Edge_weight - Combined_data$Altered_Edge_weight

#Sort by difference
Combined_data_sortededge <- Combined_data %>%
  arrange(desc(abs(Edge_diff)))

```


Plot these data
```{r}
# Filter to top 10 changed edges
top_edges <- Combined_data_sortededge %>% slice_head(n = 10)

# Plot
ggplot(top_edges, aes(x = reorder(paste(Source, Sink, sep="-"), abs(Edge_diff)), 
                      y = abs(Edge_diff), fill = SampleID)) +
  geom_col() +
  coord_flip() +
  scale_fill_brewer(palette = "Spectral") +
  labs(x = "Edge (Source - Sink)", y = "Absolute Edge Difference", 
       title = "Top 10 Most Altered Edges") +
  theme_minimal() +
  theme(axis.text.x = element_text(family = "sans", color = "black", size = 14),
        axis.text.y = element_text(family = "sans", color = "black", size = 14),
        axis.title.x = element_text(family = "sans", color = "black", size = 14),
        axis.title.y = element_text(family = "sans", color = "black", size = 14))
```





Save the data
```{r}
saveRDS(Dataset2g_Res, here("Output", "SE_Lionness", "Dataset2g_Res_100sp.rds"))

write.csv(Combined_data_sortededge, here("Output", "SE_Lionness", "Edges_100sp.csv"))
```



# 4. 10% Changes

## 4.1 - Prep the Data
```{r}
# Read in the raw data
original_data90 <- read.csv(here("Data", "GLV_SimData", "Dataset_2g",
                               "SE_50sbj100taxa_90.csv"))

original_data90 <- original_data90 %>% column_to_rownames("X") 
Dataset2g_90 <- list()
Dataset2g_90$Counts <- original_data90

```


## 4.2 - Group Net
Run SE to get a network for all the data

```{r}
set.seed(12345)

# Set seed
pseed <- list(rep.num=50, seed=10010)

# Run the SpiecEasi function
net_data <- spiec.easi(
    data.matrix(original_data90),
    method = "glasso",
    sel.criterion = "bstars", 
    lambda.min.ratio = 0.01,
    nlambda = 200,
    pulsar.select = TRUE,
    pulsar.params = pseed,  
    icov.select = "pulsar.select",
    icov.select.params = "pulsar.params", 
    lambda.log = TRUE   
  )
  

# Store the results in Dataset2g_90
Dataset2g_90$SpeicEasi_Group <- net_data
```


## 4.3 - Individual Net

Run LIONESS to get the individual networks of the unaltered  data
```{r}
# Prep the data
################################################
# List to store individual networks
Ind_Nets_90 <- list()

# Overall network
net_all <- as.matrix(SpiecEasi::getOptCov(Dataset2g_90[["SpeicEasi_Group"]]))


# Individualized Networks
################################################
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)

# Loop
for (j in 1:nrow(original_data90)) {
      samplename <- rownames(original_data90[j, ])
      nsamples <- nrow(original_data90)
      net_single <- spiec.easi(data.matrix(original_data90[-j,]),
                                method = "glasso",
                                sel.criterion = "bstars", 
                                lambda.min.ratio = 0.01,
                                nlambda = 200,
                                pulsar.select = TRUE,
                                pulsar.params = pseed,  
                                icov.select = "pulsar.select",
                                icov.select.params = "pulsar.params", 
                                lambda.log = TRUE)
      net_single_cov <- as.matrix(getOptCov(net_single))
      net_name <- samplename
      Ind_Nets_90[[net_name]] <- nsamples * (net_all -
                                    net_single_cov) + net_single_cov
      colnames(Ind_Nets_90[[net_name]]) <- rownames(Ind_Nets_90[[net_name]]) <- colnames(original_data90)
    }

# Store the results 
Dataset2g_90$Ind_Nets_90 <- Ind_Nets_90
```




Convert altered networks to an edge list
```{r}
# Create an empty edge list
##########################################
Ind_Edge_90_Table <- list()

# loop through and extract edges
##########################################
for (i in seq_along(Dataset2g_90[["Ind_Nets_90"]])) {
    cov_matrix <- Dataset2g_90[["Ind_Nets_90"]][[i]]
    network_name <- names(Dataset2g_90[["Ind_Nets_90"]][i])
    edge_table_name <- paste0(network_name)
    diagonal_indices <- 1:nrow(cov_matrix)
    cov_matrix[cbind(diagonal_indices, diagonal_indices)] <- 0
    cov_matrix[lower.tri(cov_matrix)] <- 0
    adj <- as.data.frame(as.table(cov_matrix))
    colnames(adj) <- c("Source", "Sink", "Edge_weight")
    adj <- adj[adj$Edge_weight != 0, ]
    adj$SampleID <- network_name
    Ind_Edge_90_Table[[edge_table_name]] <- adj
}

Dataset2g_90$Ind_Edge_90_Table <- Ind_Edge_90_Table
Merged_Edge_Table <- do.call(rbind, Ind_Edge_90_Table)


saveRDS(Dataset2g_90, here("Output", "SE_Lionness", "Dataset2g_Res_100sp_90.rds"))
```


Reformat the table
```{r}
# Read in the object
#Dataset2g_90 <- readRDS(here("Output", "SE_Lionness","Dataset2g_Res_100sp_90.rds"))
# rename Edge_weight to Altered_Edge_weight
Altered90_data <- Merged_Edge_Table %>%
  rename("Edge_weight"="M90_Edge_weight")

```

Add in the index data
```{r}
# Read in the index Data
Subj50_Taxa100_90_indices <- read.csv(here("Data", "GLV_SimData", "Dataset_2g",
                                           "50sbj100taxa_90indices.csv")) 

# Mutate
Subj50_Taxa100_90_indices <- Subj50_Taxa100_90_indices %>% 
                                rename("SubjectID" = "SampleID") %>%
                                rename("Source" = "Changed_species") %>%
                                dplyr::select(-X) 
 
Subj50_Taxa100_90_indices$SampleID  <- gsub("Sbj","", Subj50_Taxa100_90_indices$SampleID)
Subj50_Taxa100_90_indices$SampleID  <- gsub("_Time.*","", Subj50_Taxa100_90_indices$SampleID)

# Merge the altered edge data with the indices
Merged_90s_data <- merge(Altered90_data, Subj50_Taxa100_90_indices, 
                         by = c("SampleID"), all=TRUE)

Merged_90s_data <- Merged_90s_data %>% mutate(Edge_changed = case_when(Source == Changed_species |
                                                                       Sink == Changed_species ~ "Altered",
                                                                       TRUE ~ "Not-Altered"))

```



## 4.4 - Compare

Combine with the normal unchanged networks data
```{r}
# Merge normal data with data that is 90% consistent per subject
######################################
#update sample names
Unaltered_edges2 <- Unaltered_edges
Unaltered_edges2$SampleID  <- gsub("Sbj","", Unaltered_edges2$SampleID)
Unaltered_edges2$SampleID  <- gsub("_Time.*","", Unaltered_edges2$SampleID)

Final_edges <- merge(Unaltered_edges2, Merged_90s_data, 
                          by= c("Source", "Sink", "SampleID"))

# Check the differences
mean(Final_edges$Edge_weight)
mean(Final_edges$M90_Edge_weight)
max(Final_edges$Edge_weight)
max(Final_edges$M90_Edge_weight)
min(Final_edges$M90_Edge_weight)
min(Final_edges$Edge_weight)
sd(Final_edges$M90_Edge_weight)
sd(Final_edges$Edge_weight)

# Edge difference
Final_edges$Edge_diff <- Final_edges$Edge_weight - Final_edges$M90_Edge_weight

# Add rank
Final_edges$Original_Rank <- rank(-abs(Final_edges$Edge_weight), ties.method = "min")
Final_edges$Altered_Rank <- rank(-abs(Final_edges$M90_Edge_weight), ties.method = "min")
Final_edges$Rank_diff <- Final_edges$Original_Rank - Final_edges$Altered_Rank


```


Plot Top 10% Edge Differences
```{r}
# Plot top 10% of edge changes by edge diff
###################################################
#Sort by difference
Final_edges <- Final_edges %>%
  arrange(desc(abs(Edge_diff)))

Total_e_changed <- Final_edges %>%
  filter(Edge_changed == "Altered") %>%
  nrow()

Total_e_unchanged <- Final_edges %>%
  filter(Edge_changed == "Not-Altered") %>%
  nrow()

# Filter to top 10%
Final_edges_10 <- Final_edges %>% slice_head(n = round(nrow(Final_edges)*.1)) %>% 
                mutate(Edge_changed = as.character(Edge_changed))

Final_edges_10_summary <- Final_edges_10 %>%
  dplyr::count(Edge_changed) %>%
  mutate(overall_percentage = (n / sum(n))) %>%
  mutate(scaled_percentage = 1)

Final_edges_10_summary[1,4] <- Final_edges_10_summary$n[1] / Total_e_changed

Final_edges_10_summary[2,4] <- Final_edges_10_summary$n[2] / Total_e_unchanged


# Plot those data
ggplot(Final_edges_10_summary, aes(x = Edge_changed, y = overall_percentage, fill = Edge_changed)) +
  geom_bar(stat = "identity") +
  labs(x = "Species in Edge Changed", y = "Proportion of top 10% edges", 
       title = "Overall") +
  ylim(0,1) +
  theme_classic() +
  theme(legend.position = "none")  +
    scale_fill_manual(name = "Edge Type", 
                     values = c("Altered" = "darkgreen", 
                                "Not-Altered" = "grey"))  +
  theme(axis.text.x = element_text(family = "sans", color = "black", size = 14),
        axis.text.y = element_text(family = "sans", color = "black", size = 14),
        axis.title.x = element_text(family = "sans", color = "black", size = 14),
        axis.title.y = element_text(family = "sans", color = "black", size = 14))


# Plot those data
ggplot(Final_edges_10_summary, aes(x = Edge_changed, y = scaled_percentage, fill = Edge_changed)) +
  geom_bar(stat = "identity") +
  labs(x = "Species in Edge Changed", y = "Proportion of top 10% edges", 
       title = "Scaled %") +
  ylim(0,1) +
  theme_classic() +
  theme(legend.position = "none") +
    scale_fill_manual(name = "Edge Type", 
                     values = c("Altered" = "darkgreen", 
                                "Not-Altered" = "grey"))  +
  theme(axis.text.x = element_text(family = "sans", color = "black", size = 14),
        axis.text.y = element_text(family = "sans", color = "black", size = 14),
        axis.title.x = element_text(family = "sans", color = "black", size = 14),
        axis.title.y = element_text(family = "sans", color = "black", size = 14))
```











