---
title: "R Notebook"
output: html_notebook
---

# 1. Load Library
*** 
Packages required to run the script
```{r}
library(tidyverse)
library(igraph)
library(NBZIMM)
library(SpiecEasi)
library(LIMON)
library(here)
library(lme4)
library(Matrix)
library(MASS)
library(matrixcalc)
library(gridExtra)
library(devtools)
library(reshape2)
library(ggpubr)
library(broom)
library(ggnewscale)
library(coin)
library(parallel)
library(graph)
library(phyloseq)
library(metagMisc)
library(ggeffects)
library(car)
```



# 2. Data
*** 

## 2.1 - Background
In this script, we will be running the full LIMON pipeline on an infant dataset published in 2019. The data were pulled from this summarized article https://www.nature.com/articles/s41522-022-00345-5/figures/1, specifically reference the study by He et al 2019 


He, Xuan, et al. "Fecal microbiome and metabolome of infants fed bovine MFGM supplemented formula or standard formula with breast-fed infants as reference: a randomized controlled trial." Scientific reports 9.1 (2019): 1-14.


## 2.2 - Data set up

Load the Metadata
```{r}
data("He_2019_metadata")
```

filter down and create some binarys
```{r}
metadata <- He_2019_metadata %>% 
  dplyr::select(Subject, X, Study.Group, Age, Gender, diet) %>%
  group_by(Subject) %>%
  # Filter to those with at least three visits
  filter(n() >= 4) %>%
  ungroup() %>%
  column_to_rownames("X") %>%
  # add a binary column for Gender
  mutate(Gender_binary = case_when(
    Gender == "Male" ~ 1,
    Gender == "Female" ~ 0)) %>%
  # add a binary column for Diet (Formula or Breastmilk)
  mutate(Diet_binary = case_when(diet == "Standard infant formula" ~ 0,
                                 diet == "Experimental infant formula" ~ 0,
                                 diet == "Breast milk" ~ 1))

```



Microbial Data
```{r}
# Microbial Counts
############################################################
data("He_2019_data")

# Mutate column names to keep only the string after the last ";"
colnames(He_2019_data) <- colnames(He_2019_data) %>% 
  str_extract("([^;]+)$")
colnames(He_2019_data) <- make.names(colnames(He_2019_data), unique = TRUE)
He_2019_data <- He_2019_data %>% column_to_rownames("X")

# Ensure no columns sum to 0
raw_counts <- He_2019_data %>%
  dplyr::select(where(~ sum(.) > 0))

```

```{r}
# Look at the data
hist(as.matrix(raw_counts), breaks = 100)
```

Check Mean and standard deviation of the data
```{r}
# Stats
median(as.matrix(raw_counts))
mean(as.matrix(raw_counts))
sd(as.matrix(raw_counts))

# Zeros
percent_zeros <- sum(raw_counts== 0) / (nrow(raw_counts) * ncol(raw_counts)) * 100
percent_zeros
```
75% of the data is 0's, and the standard deviation is >> then the mean so it follows a zero-inflated negative binomial distribution


Check some of the features abundance by gender
```{r}
# Step 1 - Merge metadata with count table
################################################
common_samples <- intersect(rownames(raw_counts), rownames(metadata))
raw_counts <- raw_counts[common_samples, ]
metadata <- metadata[common_samples, ]

# Merge
counts_meta <- merge(metadata, as.matrix(raw_counts), by = 0, all=TRUE)
counts_meta <- counts_meta %>% column_to_rownames("Row.names")

# Step 2 - Mean Abundance by diet
################################################
mean_data <- counts_meta %>%
  group_by(diet = counts_meta[[5]]) %>%
  summarise(across(8:81, mean, na.rm = TRUE)) %>%
  pivot_longer(-diet, names_to = "Variable", values_to = "Mean")

# Plot the mean values by diet
ggplot(mean_data, aes(x = Variable, y = Mean, fill = diet)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkviolet","steelblue", "darkgreen")) +
  labs(x = "Genus", y = "Mean Abundance Count", title = "Mean Counts by Diet") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Step 2 - Mean Abundance by gender
################################################
mean_data <- counts_meta %>%
  group_by(Gender = counts_meta[[4]]) %>%
  summarise(across(8:81, mean, na.rm = TRUE)) %>%
  pivot_longer(-Gender, names_to = "Variable", values_to = "Mean")

# Plot the mean values by gender
ggplot(mean_data, aes(x = Variable, y = Mean, fill = as.factor(Gender))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkviolet","darkgreen")) +
  labs(x = "Genus", y = "Mean Abundance Count", title = "Mean Counts by Gender") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```



CSS normalize the counts
```{r}
# Create the OTU table 
otu_table <- otu_table(raw_counts, taxa_are_rows = FALSE)
sample_data <- sample_data(metadata)

# CSS norm in Phyloseq
physeq <- phyloseq(otu_table, sample_data)
ps.css <- phyloseq_transform_css(physeq, norm= TRUE, log = TRUE)

normalized_counts <- round(as.data.frame(t(ps.css@otu_table)))
```


Check some of the features abundance by gender
```{r}
# Step 1 - Merge metadata with count table
################################################
common_samples <- intersect(rownames(normalized_counts), rownames(metadata))
normalized_counts <- normalized_counts[common_samples, ]
metadata <- metadata[common_samples, ]

# Merge
counts_meta <- merge(metadata, as.matrix(normalized_counts), by = 0, all=TRUE)
counts_meta <- counts_meta %>% column_to_rownames("Row.names")

# Step 2 - Mean Abundance by diet
################################################
mean_data <- counts_meta %>%
  group_by(diet = counts_meta[[5]]) %>%
  summarise(across(8:81, mean, na.rm = TRUE)) %>%
  pivot_longer(-diet, names_to = "Variable", values_to = "Mean")

# Plot the mean values by diet
ggplot(mean_data, aes(x = Variable, y = Mean, fill = diet)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkviolet","steelblue", "darkgreen")) +
  labs(x = "Genus", y = "Mean Abundance Count", title = "Mean Counts by Diet") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Step 2 - Mean Abundance by gender
################################################
mean_data <- counts_meta %>%
  group_by(Gender = counts_meta[[4]]) %>%
  summarise(across(8:81, mean, na.rm = TRUE)) %>%
  pivot_longer(-Gender, names_to = "Variable", values_to = "Mean")

# Plot the mean values by gender
ggplot(mean_data, aes(x = Variable, y = Mean, fill = as.factor(Gender))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkviolet","darkgreen")) +
  labs(x = "Genus", y = "Mean Abundance Count", title = "Mean Counts by Gender") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```


# 3. LIMON
*** 

```{r}
# Ensure sample counts and metadata are in the same order
common_samples <- intersect(rownames(normalized_counts), rownames(metadata))
normalized_counts <- normalized_counts[common_samples, ]
metadata <- metadata[common_samples, ]


# Make LIMON object
L_obj <- LIMON_Obj(Counts = as.matrix(normalized_counts), 
                           SampleData = metadata)
```


```{r}
# Set seed
set.seed(12345)
# Fit the distribution/remove covariates
#########################################
L_obj2 <- LIMON_DistrFit(Obj = L_obj, 
                           Time = "Age", 
                           Subject = "Subject", 
                           Covariates = c("Gender_binary"),
                           model = "Gender_binary",
                           distribution = "GLMM.Z")

```




```{r}
# Check the data cleaning
################################################
cleaned <- merge(metadata, L_obj2[["Corrected_Counts"]], by = 0)
cleaned <- cleaned %>% column_to_rownames("Row.names")

# Step 2 - Mean Abundance by gender
################################################
mean_data <- cleaned %>%
  group_by(Gender = cleaned[[4]]) %>%
  summarise(across(8:81, mean, na.rm = TRUE)) %>%
  pivot_longer(-Gender, names_to = "Variable", values_to = "Mean")

# Plot the mean values by gender
ggplot(mean_data, aes(x = Variable, y = Mean, fill = as.factor(Gender))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkviolet","darkgreen")) +
  labs(x = "Genus", y = "Mean Abundance Count", title = "Gender Corrected") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Networks per timepoint
```{r}
set.seed(12345)
# SPIEC-EASI per time
# Set seed
pseed <- list(rep.num=50, seed=10010)


#infer network
L_obj3 <- LIMON_NetInf_Time(Obj = L_obj2, 
                                         method = "glasso", 
                                         sel.criterion = "bstars",
                                         lambda.min.ratio = 0.01,
                                         pulsar.select=TRUE, 
                                         pulsar.params=pseed,
                                         nlambda = 200)



# Print Networks
L_obj4 <- LIMON_Edges_Networks(L_obj3, threshold = 0.02, vertex.size = 3, 
                                       vertex.label.cex = 8, vertex.label.color = "black")
```


Individualized Networks
```{r}
# Set seed
set.seed(12345)
pseed <- list(rep.num=50, seed=10010)


# individual Networks
L_obj6 <- LIMON_IndNet(Obj = L_obj3, method = "glasso", 
                                         sel.criterion = "bstars",
                                         lambda.min.ratio = 0.01,
                                         pulsar.params=pseed,
                                         nlambda = 200)


# Save the object
saveRDS(L_obj6, here("Output","He_2019", "HE2019_LIMON.rds"))
```





```{r}
# option to read the object back in
L_obj6 <- readRDS(here("Output","He_2019", "HE2019_LIMON.rds"))

# Extract edges and centralities
L_obj7<- LIMON_IndEdges(L_obj6, threshold = 0.02)
L_obj8 <- LIMON_Centralities(L_obj7, threshold = 0.02)
```




LIMON will run a statistical model with per each unique interaction at each time point with that interactions edge weight as the predictor, and a user specified dependent variable as the outcome. The current model options are "lm" for a linear regression (for a continuous outcome variable), "logit" for a logistic regression (for a binary outcome variable only), and "multinom", multinomial model (for 3 or more categories). There is a minimum number of observations required for the different models to run; 10 for lm, 20 for logit, and 30 for multinom. If there are not enough observations for that interaction at that timepoint, a message will be returned to the user. 

The defualt is to not filter the models by signficance. Users can set a p-value or adjusted p-value threshold to limit the number of models returned. This is also useful for printing the graphs from the linear and logistic models. Users can specify if they want to print a graph of all of their interactions returned fromt he model (plot_nodes = TRUE). This will not work for the multinomial model however, we will address how to visualize those data in a bit. 

Our question of interst in this dataset is how do the different diets (standard formula; SF, experimental formula; EF, and breast milk; BF) alter microbial interactions across the first year of life (2,4,6,12 months). Lets first start by looking at the binary diet outcome we created, which was formula fed (SF or EF; 0) vs Breasf Fed (1). Run a logisitc model and print the graphs with a p-value threshold of 0.05 
```{r}
# Create a new object
L_obj9 <- L_obj8


L_obj9 <- LIMON_StatNodes(L_obj9, dependent = "Diet_binary", time = "Age", pval = 0.05,
                           method ="logit", timepoints= NULL, plot_nodes = TRUE, 
                           estimate = NULL, custom_colors = NULL, point_size = 5)
```


```{r}
# Create a new object
L_obj9 <- L_obj8


L_obj9 <- LIMON_StatNodes(L_obj9, dependent = "diet", time = "Age", pval = 0.05,
                           method ="multinom", timepoints= NULL, plot_nodes = FALSE, 
                           estimate = NULL, custom_colors = NULL, point_size = 8)

# Examine the Output
nodes_data <- L_obj9[["Significant_Interactions"]]
```


g__Parabacteroides-g__Actinomyces time 6
```{r}
Edge_Table <- L_obj8[["Merged_Edge_Table"]]
        
# list to store all of the model data in
node_results <- list()

# Filter data to a timepoint of interest        
edge_data_time_full <- Edge_Table %>% filter(Edge_Table[["Age"]] == 6)

# Filter to an interaction of interest
edge_data_time <- edge_data_time_full %>% filter(Interaction == "g__Parabacteroides-g__Actinomyces")
dependent <- "diet"

# Run the model
formula <- as.formula(paste(dependent, "~Edge_weight"))
set.seed(12345)
model <- nnet::multinom(formula, data = edge_data_time, trace = FALSE)
model_summary <- broom::tidy(model, conf.int = TRUE)
model_summary$Interaction <- "g__Parabacteroides-g__Actinomyces"
model_summary$Time_Level <- 6
model_summary$Model_Type <- "Multinomial"
model_summary$Model_SampleSize <- stats::nobs(model)
node_results[[paste("g__Parabacteroides-g__Actinomyces", 6, sep = "_")]] <- model_summary


# get probabilities
set.seed(12345)
predict(model, newdata = data.frame(Edge_weight = 2.5), type = 'probs')
set.seed(12345)
prob_effects <- ggeffect(model, terms = "Edge_weight[-2:2,by=0.5]")

# Graph probabilities
ggplot(prob_effects) +
  aes(x = x, y = predicted, fill = response.level, color = response.level) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 1/2) +
  labs(x = 'g__Parabacteroides-g__Actinomyces Edge Weight', y = 'Predicted Probability') +
  ylim(c(0, 1)) +
  theme_classic() +
  scale_color_manual(
    values = c("#bc80bd", "#ffed6f", "#ccebc5"), 
    name = "Infant Diet",
    labels = c("Breast Milk" = "Breast.milk", 
               "Experimental infant formula" = "Experimental.infant.formula", 
               "Standard infant formula" = "Standard.infant.formula") ) +
  scale_fill_manual(
    values = c("#bc80bd", "#ffed6f", "#ccebc5"), 
    name = "Infant Diet",
    labels = c("Breast Milk" = "Breast.milk", 
               "Experimental infant formula" = "Experimental.infant.formula", 
               "Standard infant formula" = "Standard.infant.formula") ) +
  theme(
    axis.text.x = element_text(color = "black", family = "Arial", size = 11), 
    axis.text.y = element_text(color = "black", family = "Arial", size = 11))


# Check model 
set.seed(12345)
car::Anova(model)
```



g__Parabacteroides-g__Actinomyces time 12
```{r}
Edge_Table <- L_obj8[["Merged_Edge_Table"]]
        
# list to store all of the model data in
node_results <- list()

# Filter data to a timepoint of interest        
edge_data_time_full <- Edge_Table %>% filter(Edge_Table[["Age"]] == 12)

# Filter to an interaction of interest
edge_data_time <- edge_data_time_full %>% filter(Interaction == "g__Parabacteroides-g__Actinomyces")
dependent <- "diet"

# Run the model
formula <- as.formula(paste(dependent, "~Edge_weight"))
set.seed(12345)
model <- nnet::multinom(formula, data = edge_data_time, trace = FALSE)
model_summary <- broom::tidy(model, conf.int = TRUE)
model_summary$Interaction <- "g__Parabacteroides-g__Actinomyces"
model_summary$Time_Level <- 12
model_summary$Model_Type <- "Multinomial"
model_summary$Model_SampleSize <- stats::nobs(model)
node_results[[paste("g__Parabacteroides-g__Actinomyces", 12, sep = "_")]] <- model_summary


# get probabilities
set.seed(12345)
predict(model, newdata = data.frame(Edge_weight = 2.5), type = 'probs')
set.seed(12345)
prob_effects <- ggeffect(model, terms = "Edge_weight[-2:2,by=0.5]")

# Graph probabilities
ggplot(prob_effects) +
  aes(x = x, y = predicted, fill = response.level, color = response.level) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 1/2) +
  labs(x = 'g__Parabacteroides-g__Actinomyces Edge Weight', y = 'Predicted Probability') +
  ylim(c(0, 1)) +
  theme_classic() +
  scale_color_manual(
    values = c("#bc80bd", "#ffed6f", "#ccebc5"), 
    name = "Infant Diet",
    labels = c("Breast Milk" = "Breast.milk", 
               "Experimental infant formula" = "Experimental.infant.formula", 
               "Standard infant formula" = "Standard.infant.formula") ) +
  scale_fill_manual(
    values = c("#bc80bd", "#ffed6f", "#ccebc5"), 
    name = "Infant Diet",
    labels = c("Breast Milk" = "Breast.milk", 
               "Experimental infant formula" = "Experimental.infant.formula", 
               "Standard infant formula" = "Standard.infant.formula") ) +
  theme(
    axis.text.x = element_text(color = "black", family = "Arial", size = 11), 
    axis.text.y = element_text(color = "black", family = "Arial", size = 11))

set.seed(12345)
car::Anova(model)
```



