---
title: "LIMON Data Simulation - Dataset 1"
output: html_notebook
---

# Part 1: Load Library
***
```{r}
library(tidyverse)
library(igraph)
library(NBZIMM)
library(SpiecEasi)
library(LIMON)
library(here)
library(lme4)
library(Matrix)
library(tscount)
library(patchwork)
library(MASS)
library(matrixcalc)
library(gridExtra)
library(devtools)
library(miaSim)
library(reshape2)
```


# 2. Simulate counts using Generalized Lotka Volterra
***
Github for miaSim https://github.com/microbiome/miaSim 


Simulate GLV for 50 individuals, 50 species, 10 timepoints, dense network


__Simulate Counts__  
Simulate GLV for 50 individuals, 50 species, 10 time points, dense n
```{r}
# Set seed
set.seed(12345)

# Step 1 Run GLV for n number of subject and timepoints
###############################################################################

# Generate interactions from uniform distribution
A_uniform <- randomA(
    n_species = 50,
    #scale_off_diagonal = 1,
    diagonal = -1.0,
    connectance = 0.9)

hist(A_uniform, breaks = 100)

# Create an empty list to store the count tables for each subject
count_tables <- list()


# Loop through 50 subjects and generate count tables for each
for (i in 1:50) {
  # Set seed
  set.seed(12345+i)
 # Generalized Lotka-Volterra (gLV)
  tse_glv <- simulateGLV(n_species = 50,
                       #x0 = NULL,
                       A = A_uniform,
                       t_start = 0, 
                       t_store = 10,
                       stochastic = FALSE,
                       norm = FALSE,
                       error_variance = 0.01)
  
  # Get the count table
  sim_data <- tse_glv@assays@data@listData[["counts"]]
  
  # Store the count table in the list
  count_tables[[i]] <- t(sim_data)
}

# Step 2 - Merge together
###############################################################################
# Combine all count tables into one data frame
combined_count_table <- do.call(rbind, count_tables)

# Rename the rownames based on the count table number
rownames(combined_count_table) <- paste0("Sbj", rep(1:50, each = nrow(count_tables[[1]])), "_Time", 1:10)

```






__Create Fake Metadata__  
1. Sex (M or F, 50/50 Ratio)
2. Age - sample from between 18 and 45
3. BMI - sample between 18 and 35

Make Metadata and merge with the count data
```{r}
# Df 1 is Metadata
########################################################
meta_data <-  expand.grid(Time = 1:10,ID = 1:50)
rownames(meta_data) <- rownames(combined_count_table)
# Set seed
set.seed(12345)
meta_data$Sex <- rep(c(0, 1), each = 50)
# Set seed
set.seed(12345)
meta_data$Age <- rep(sample(18:45, 50, replace = TRUE), each = 10)
# Set seed
set.seed(12345)
meta_data$BMI <- rep(sample(18:35, 50, replace = TRUE), each = 10)

# Center the continuous variables
meta_data$Age <- meta_data$Age - mean(meta_data$Age)
meta_data$BMI <- meta_data$BMI - mean(meta_data$BMI)


# Df 2 is Metadata merged with Counts
########################################################
#Round off and increase
combined_count_table <- as.data.frame(combined_count_table + abs(min(combined_count_table)))
combined_count_table <- (combined_count_table)*10
meta_counts <- base::merge(meta_data, combined_count_table, by ="row.names", all = TRUE)
meta_counts <- column_to_rownames(meta_counts, "Row.names")
```




Graphs to Check
```{r}
# Individual Species Plots
########################################################
# Pivot to long data
count_long <- tidyr::pivot_longer(meta_counts, cols = starts_with("sp"), names_to = "Species")

# Plot the data
count_long %>%
  ggplot(aes(x = Time, y = value, colour = as.factor(ID),
             group = as.factor(ID), linetype = as.factor(ID))) +
  geom_line() + 
  geom_point() +
  geom_jitter() +
  ylab("Count") +
  labs(linetype = "ID", color = "ID") +
  facet_wrap(~ Species) +  # Create a panel for each species
  theme(legend.position = "none") +
  ggtitle("Time Series of N=10, 0s")


# Plot only two
########################################################
# Pivot to long data
metacounts_filt <- meta_counts[,c(1:7)]
count_long <- tidyr::pivot_longer(metacounts_filt, cols = starts_with("sp"), names_to = "Species")

# Plot the data
count_long %>%
  ggplot(aes(x = Time, y = value, colour = as.factor(ID),
             group = as.factor(ID), linetype = as.factor(ID))) +
  geom_line() + 
  geom_point() +
  geom_jitter() +
  ylab("Count") +
  labs(linetype = "ID", color = "ID") +
  facet_wrap(~ Species) +  # Create a panel for each species
  theme_linedraw() +
  theme(axis.text.x = element_text(family = "arial",color = "black", size = 14),
        axis.text.y = element_text(family = "arial",color = "black", size = 14),
        axis.title.x = element_text(family = "arial",color = "black", size = 14),
        axis.title.y = element_text(family = "arial",color = "black", size = 14),
        legend.position = "none",
        strip.text = element_text(face = "bold", family = "arial", color = "white", size = 15))


# Distribution of counts
########################################################
hist(as.matrix(combined_count_table), breaks = 100, main = "Distribution of GLV Data", xlab = "Counts")

```



Add in biological covariates
```{r}
# Set seed
set.seed(12345)
# Addin covariates
########################################################################################
# Set up new dataframe
Long_data_new <- meta_counts

# Taxa 21 - 30 will have Sex effect
for (i in 26:35) {
  error <- rnorm(nrow(Long_data_new), mean = 1, sd = 0.6)
  Long_data_new[, i] <- Long_data_new[, i] + 8 * Long_data_new$Sex + error
}

# round the counts to bring them back up to 0
########################################################################################

# Add the minimum value to bring everything up to at least 0
count_table1 <- Long_data_new[,6:55]


# scale to positive and make larger
count_table1 <- count_table1 + abs(min(count_table1))
count_table1 <- round(count_table1*10)

#change Long_data_new
Long_data_new[,6:55] <- count_table1

# Correlation matrix
hist(as.matrix(Long_data_new[,6:55]), breaks = 100, main = "Distribution of GLV Data +cov", xlab = "Counts")
cor_raw1 <- cor((Long_data_new[,6:55]), method = "spearman")
heatmap(cor_raw1, Colv = NA, Rowv = NA, main = "Correlation of data + covariates")
```



Now Add in 0s
```{r}
# Set up new dataframe
predata_0 <- count_table1

# Add the 0s back in
########################################################################################


# Step 1: Calculate total counts for each column
total_counts <- colSums(predata_0)


# Step 2: Create probability gradient
gradient <- seq(0.5, 0.2, length.out = ncol(predata_0))

# Step 3: Make the probability gradient inverse to total counts (ie higher total value, lower proportion of 0s)
total_counts <- total_counts[order(total_counts)]
gradient <- gradient[order(-gradient)]

# Step 4 & 5: Generate random numbers and set counts to 0 based on probability gradient

for (i in seq_along(total_counts)) {
  prob <- gradient[i]
  # Calculate number of 0s to add based on probability
  num_zeros <- sum(runif(nrow(predata_0)) <= prob)
  # Randomly select rows to set to 0
  # Set seed
  set.seed(12345+i)
  rows_to_zero <- sample(nrow(predata_0), num_zeros)
  # Set counts to 0
  predata_0[rows_to_zero, i] <- 0
}

# merge with metadata for plotting
zero_data1 <- merge(meta_data, predata_0, by = 0)
zero_data1 <- column_to_rownames(zero_data1, "Row.names")
#round the counts
zero_data1[,6:55] <- round(zero_data1[,6:55])
```


Graphs to Check
```{r}
# Individual Species Plots
########################################################
# Pivot to long data
count_long <- tidyr::pivot_longer(zero_data1, cols = starts_with("sp"), names_to = "Species")

# Plot the data
count_long %>%
  ggplot(aes(x = Time, y = value, colour = as.factor(ID),
             group = as.factor(ID), linetype = as.factor(ID))) +
  geom_line() + 
  geom_point() +
  geom_jitter() +
  ylab("Count") +
  labs(linetype = "ID", color = "ID") +
  facet_wrap(~ Species) +  # Create a panel for each species
  theme(legend.position = "none") +
  ggtitle("Time Series of N=10, 0s")


# Distribution of counts
########################################################
hist(as.matrix(zero_data1[,6:55]), breaks = 100, main = "Distribution of GLV Data", xlab = "Counts")

# Correlation matrix
cor_raw1 <- cor((zero_data1[,6:55]), method = "spearman")
heatmap(cor_raw1, Colv = NA, Rowv = NA, main = "Correlation of 0 inflated no covariates")
```


# 3. Save the data

Save the Counts
```{r}
# Make more like counts
meta_counts[,6:55] <- round(meta_counts[,6:55]*10)
write.csv(meta_counts, here("Data","GLV_SimData", "Dataset_1", "GLV.csv"))
write.csv(Long_data_new, here("Data","GLV_SimData", "Dataset_1","GLV_Cov.csv"))
write.csv(zero_data1, here("Data","GLV_SimData", "Dataset_1","GLV_CovZero.csv"))
```

